---
title: "3. DA_Cluster"
author: "Xin Tan, 583833"
date: "2019/3/11"
output:
  word_document: default
  html_document:
    df_print: paged
---


```{r eval=FALSE, include=FALSE}
# Test 1
library(readxl)
dat <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "AttributeRatingsStacked")
dat<-as.data.frame(dat)


dat[is.na(dat)] <- 0

attribute.rating<- aggregate(dat[,-c(1,2)], by=list(dat$Person),mean, na.rm=F) 
#attribute.rating[is.na(attribute.rating)] <- 0
colnames(attribute.rating)[1] <- "Person"



# Test 2 4 attribute

library(psych)
apply(dat,2,sd)
# replace with 0。表示考虑na时的方差。用这个。
# Calorie,rich,healthful,sweet -> rich (1.563614)
# Commercial, image, chocolaty -> commercial (1.569881)
# Creamy, crunchy, addiction, wrapping -> addiction (1.524589)
# Accessible, handy -> accessible (1.427612)

# remove Na. 这个表示不考虑na的时候已有数据的方差。
apply(dat,2,sd,na.rm=T)
# Calorie,rich,healthful,sweet -> rich (1.0659400)
# Commercial, image, chocolaty -> commercial (1.1571558 )
# Creamy, crunchy, addiction, wrapping -> crunchy(1.3878090)
# Accessible, handy -> handy (0.8890739)

# 构建数据集


# by mean
library(readxl)
dat <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "AttributeRatingsStacked")
dat<-as.data.frame(dat)
dat[is.na(dat)] <- 0

dat<- dat[c("Person","Product","rich","commercial","addiction","accessible","crunchy","creamy")]
attribute.rating<- aggregate(dat[,-c(1,2)], by=list(dat$Person),mean, na.rm=T)
colnames(attribute.rating)[1] <- "Person"




```


In the cluster, NA means the respondents have no experence about the rating objective, which should not be imputate with another number, otherwise we will lose this important information. Here I will replace the NA with zero, on the  one hand, the information has been kept, on the other hand, the orienginal scale and the distribution of data is not influenced too much.

```{r}

# prepare for the attribute table 结果虽好，但有问题：一旦有个没填，整个attribute都是0
library(readxl)
dat <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "AttributeRatingsStacked")
dat<-as.data.frame(dat)

# attribute.rating<- aggregate(dat[,-c(1,2)], by=list(dat$Person),mean, na.rm=F)
attribute.rating<- aggregate(dat[,-c(1,2)], by=list(dat$Person),mean, na.rm=T)
attribute.rating[is.na(attribute.rating)] <- 0
colnames(attribute.rating)[1] <- "Person"

```





# Cluster only use attribute
```{r eval=FALSE, fig.height=15, fig.width=15, include=FALSE}
# Calinski and Harabasz (suggest 2 cluster)
library("NbClust")

NbClust(attribute.rating[-1], method="complete")
NbClust(attribute.rating[-1], method="ward.D2")
NbClust(attribute.rating[-1], method="kmeans")




#H-cluster
library(cluster)
d <- dist(attribute.rating[-c(1)])
tmp.cluster <-hclust(d, method = "complete") 
tmp.cluster <-hclust(d, method = "ward.D2") 


plot(as.dendrogram(tmp.cluster),main = "Ward Cluster Dendrogram", horiz = FALSE)
rect.hclust(tmp.cluster , k = 3, border = "red")

```


```{r}

# Silhouette of H-cluster (2 clusters is better)
cl1  <- hclust(d, method="ward.D2")
memb2 <- cutree(cl1, 4)
memb3 <- cutree(cl1, 3)
#
library("cluster")
par(mfcol=c(2,2))
plot(attribute.rating, col=memb2)
s2 <- silhouette(memb2, d)
plot(s2, col=1:4, border=NA)
plot(attribute.rating, col=memb3)
s3 <- silhouette(memb3, d)
plot(s3, col=1:3, border=NA)




# h-cluster visualization
cl1  <- hclust(d, method="complete")
memb2 <- cutree(cl1, 2)
memb3 <- cutree(cl1, 3)
#
library("cluster")
par(mfcol=c(1,2))
clusplot(attribute.rating, memb2, col.p=memb2)
clusplot(attribute.rating, memb3, col.p=memb3)





library("cluster")
library("mlbench") # for Boston Housing data

# prepare data


# Diana (好像没什么用)
cl  <- diana(attribute.rating[-c(1)])
hcl <- cutree(as.hclust(cl), k = 3)
par(mfrow=c(4,4))
plot(attribute.rating$Person, attribute.rating$crunchy, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$creamy, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$sweet, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$chocolaty, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$healthful, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$calorie, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$rich, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$addiction, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$accessible, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$handy, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$wrapping, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$image, col=hcl, pch=19, cex=0.5)
plot(attribute.rating$Person, attribute.rating$commercial, col=hcl, pch=19, cex=0.5)

```

```{r}
# -------------------------------- kmeans
library(cluster)
library(NbClust)

# Calinski and Harabasz (suggest 2 cluster)
NbClust(attribute.rating[-1], method="kmeans")

# set.seed(3) #78.43% plot
set.seed(2)

seg.k <- kmeans(attribute.rating[-c(1)], centers= 3)

attribute.rating$kmeans <- seg.k$cluster
clusplot(attribute.rating[-c(1)], seg.k$cluster, color=TRUE, shade=TRUE,labels= , lines=0, main="K-means cluster")


table(seg.k$cluster)
table(seg.k$cluster)/50




# fit statistics
seg.k$betweenss
seg.k$totss
seg.k$betweenss/seg.k$totss






# kmeans Silhouette (不用)

library("cluster")
d <- dist(attribute.rating[-c(1)])
set.seed(2)
seg.k <- kmeans(attribute.rating[-c(1)] , centers= 2)
seg.k 
memb2 <- seg.k$cluster

s2 <- silhouette(memb2, d)
plot(s2, col=1:2, border=NA, main = "Silhouette Plot of the Kmeans Cluster")

# add kmeans in the table
library(readxl)
socio<-read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "info_xin")
ab <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "attri_xin")
socio<-as.data.frame(socio)
ab <- as.data.frame(ab)

socio$kmeans <- seg.k$cluster
bars<-merge(ab,socio,by="Person")



write.csv(bars, file = "kmeans3.csv")
# Mosaic

plot(table(socio$Gender, socio$ChocoFrequency, socio$Sport), las =2, xlab = "choco frequency", ylab = "sport frequency") # 女的比男的爱吃巧克力（结论能不能乱写）

plot(table(socio$kmeans, socio$ChocoFrequency, socio$Sport), las =2, xlab = "choco frequency", ylab = "sport frequency") # 分类为爱吃巧克力和不爱的

# 巧克力的消费程度
plot(table(socio$kmeans,socio$ChocoFrequency_Num))

# 运动
plot(table(socio$kmeans,socio$Sport_Num), las =2)

# brand
brand <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "Direct Preference Rating")
brand$kmeans <- seg.k$cluster

seg.summ <- function (data, groups) 
{aggregate (data , list(groups), function (x) mean(as.numeric(x)))}

seg.summ(brand, brand$kmeans)

# 每个人对attribute的打分, omit NA
dat <- read_excel("Data_Chocolate_allinterviews.xlsx", sheet = "AttributeRatingsStacked")
id.attr<- aggregate(dat,by=list(dat$Person),mean, na.rm=T)

id.attr$kmeans<- seg.k$cluster

tmp <- aggregate(id.attr, by=list(id.attr$kmeans),mean, na.rm=T);tmp

#Chernoff faces
library("MASS")
library("aplpack")

tmp <- tmp[-c(1:4,18)]
x <- scale(tmp)
faces(x, scale=F)
#
aggregate(bars, by=list(bars$Occupation),mean, na.rm=T)
aggregate(brand, by=list(socio$Occupation),mean, na.rm=T) # 学生最爱kr, 上班族最爱snicker
aggregate(brand, by=list(socio$Gender),mean, na.rm=T) # 男女最爱kr，


```



```{r}
library(gmodels)
aggregate(bars[c(3:15)], by=list(bars$kmeans),sd)
aggregate(bars[c(3:15)], by=list(bars$kmeans),ci)


```








